---
title: "wine_report"
author: lana almorabah
output: html_document
date: '2022-06-16'
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "../"
    )
  })
---


  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

install.packages('ggplot2')
install.packages('dplyr')
install.packages('tidyverse')
install.packages('DataExplorer')
install.packages('corrgram')
install.packages('here')
install.packages('reactable')
here::i_am("individual-assignment-2-lana7am.Rproj")
```

## MISK-DSI assigment 2

## DATA : Wine Quality dataset 



| Fixed Acidity        | Most acids involved with wine that are fixed or nonvolatile                                                      |
|----------------------|------------------------------------------------------------------------------------------------------------------|
| Volatile Acidity     | Amount of acetic acid in wine                                                                                    |
| Citric Acid          | Can add ‘freshness’ and flavor to wines.                                                                         |
| Residual Sugar       | Amount of sugar remaining after fermentation stops                                                               |
| Chlorides            | Amount of salt in the wine                                                                                       |
| Free Sulfur Dioxide  | Free Form. Prevents microbial growth and the oxidation of wine                                                   |
| Total Sulfur Dioxide | Free and bound forms of S02; in low concentrations.                                                              |
| Density              | Closeness of density to that of water depending on the percent alcohol and sugar content.                        |
| pH                   | Describes how acidic a wine is on a scale from 0 (very acidic) to 14 (very basic)                                |
| Sulphites            | Wine additive which can contribute to sulfur dioxide gas (S02) levels. Acts as an antimicrobial and antioxidant. |
| Alcohol              | the percent alcohol content of the wine (% by volume).                                                           |
| Quality              | Reviews of wines on a 1-10 scale.                                                                                |
##EDA

first initiate and download packages 

```{r init}
library(ggplot2)
library(dplyr)
library(here)
library(reactable)
```

then load the dataset 

```{r load wine}
wine_tbl <- read.csv(here::here('data/winequality-red.csv'))
```

Interactive tables 

```{r reactable, echo=FALSE}
reactable(wine_tbl)
```



applying the glimpse() function to get an overview of our data
this function allows the features to be displayed vertically while the data can be viewed horizontally 

```{r glimpse wine}
dplyr::glimpse(wine_tbl)
```

here I've applied the summary function which is used to produce result summaries of various model fitting functions.

```{r summary, echo=FALSE}
summary(wine_tbl)
reactable(wine_tbl)
```

applying the clean_names function from the janitor package
```{r clean, echo=FALSE}
janitor::clean_names(wine_tbl)
```

to get an overview of the quality of wine in this dataset i used a pie chart graph to display the proportions of each index in the quality category

```{r piechart, echo=FALSE}
(piechart <- ggplot(wine_tbl, aes( x="",y=quality,fill=factor(quality))) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0)+
  scale_fill_manual(values = c("red", "yellow",
                               "blue", "orange",
                               "green","purple")))
```

here i've used the corrgram() function to find the correlation between all the features

```{r correlation, warning=FALSE}
corrgram::corrgram(wine_tbl, order = TRUE, upper.panel =NULL)
```


from the graph it can be observed that the features with the highest correlation are:
  - density and fixed acidity
  - fixed acidity and citric acid 
  - free sulfur dioxide and total sulfur dioxide

the previous graph shows that the is a medium-high correlation between the quality of the wine and the alcohol level, therefore i wanted to calculate the average alcohol level for wines with a quality index of 7 and above.

```{r average, warning=FALSE}
wine_tbl %>% filter(quality > 6 ) %>% pull(alcohol) %>% mean()
```

##Conclusion

from the EDA we can conclude that there isn't just one feature that effects the quality of the wine drastically.
rather we can assume that its the combination of multiple features that effects the quality.
